---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# samplingNR

`samplingNR` is an R package that allows for computing optimal allocations under anticipated nonresponse, as described in our paper.

## Installation

You can install `samplingNR` by using a command such as:

```{r, eval = FALSE}
install.packages("samplingNR_0.2.1.tar.gz")
```

## Vignette

You can learn about how to use the package in `vignette("samplingNR")`, which shows how to replicate two of the tables from our paper in an application to a post-election survey of military personnel.

## Basic usage

The main allocation function is `opt_nh_nonresp()`, which provides the exact version of our proposed optimal allocation under anticipated nonresponse. The current version assumes that the goal is to minimize the (expected) variance subject to a constraint on the total (expected) costs or invited sample size.

### Fixed total sample size

If the aim is to allocate a fixed total (invited) sample size, the basic usage is:

```
opt_nh_nonresp(
  N_h,
  phibar_h,
  S_h = NULL,
  n_total
  ...
)
```

where vectors `N_h` and `phibar_h`denote the strata population sizes, anticipated response rates, and (optionally) strata variances, respectively, and where scalar `n_total` denotes the total sample size. If `S_h` is omitted, strata variances are assumed constant across strata.

### Fixed total costs

If the aim is to allocate sample subject to a constraint on total costs, the basic usage is:

```
opt_nh_nonresp(
  N_h,
  phibar_h,
  S_h = NULL,
  cost_total,
  c_NR_h,
  tau_h
  ...
)
```

Here, `cost_total` denotes the total allowable costs, `c_NR_h` denotes the unit costs per nonrespondent (by strata), and `tau_h` denotes the ratio of the unit costs per respondent to those of nonrespondents (by strata). The arguments `c_NR_h` and `tau_h` can be specified as vectors of dimension `H` if these quantities vary by strata; alternatively, if assumed constant across strata, they can be specified as scalars.
