---
title: "Sample allocation example: 2016 PEVS-ADM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sample allocation example: 2016 PEVS-ADM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The example below shows how to use `samplingNR` to allocate a sample of 50,000 invitees, which implicitly assumes use of the constant cost per invitee scenario (as defined in our paper) and assuming equivalent strata variances across strata (i.e., $S_h = S$ for some constant $S$). The [tibble](https://tibble.tidyverse.org/) `pevs_adm_2016_rrs`, included as an example data set, contains estimated population sizes and response propensities by strata for the 2016 Post-Election Voting Survey of Active Duty Military (PEVS-ADM; FVAP, 2017a, 2017b), as was used in our paper.

## Compute allocations

### Exact optimal allocation (iterative)

We can use the function `opt_nh_nonresp()` to compute the exact (i.e., iteratively computed) version of the optimal allocation under nonresponse, as provided in our paper. We omit the optional argument `S_h`, which defaults to assuming that $S_h$ is constant across strata.

```{r example}
library(samplingNR)
library(magrittr)
library(dplyr)

#Compute exact optimum allocation for PEVS-ADM 2016 data set for n = 50k
#Assumes tau = 1 by default
pevs_optE_alloc_50k <- opt_nh_nonresp(N_h = pevs_adm_2016_rrs$Nhat_h,
                                      phibar_h = pevs_adm_2016_rrs$rr_h,
                                      n_max = 50000)

#Merge results into original data frame
pevs_adm_alloc_50k_merged <- pevs_adm_2016_rrs %>%
  dplyr::mutate(n_h_optE = c(pevs_optE_alloc_50k),
                zeta_h_optE = attr(pevs_optE_alloc_50k,"zeta_h_min_1"))

#View results
pevs_adm_alloc_50k_merged
```

### Approximately optimal allocation

The approximate version of this allocation (i.e., assuming that $\zeta_h(.) \approx 1$) can be computed using the function `opt_nh_nonresp_oneiter()`.

```{r}
pevs_optA_alloc_50k <- opt_nh_nonresp_oneiter(N_h = pevs_adm_2016_rrs$Nhat_h,
                                              phibar_h = pevs_adm_2016_rrs$rr_h,
                                              n_max = 50000)
```

### Neyman-style allocations for comparison

For comparison purposes, we can also compute Neyman allocation of invitees ($n_h^{\textrm{Ninv}}\propto N_h S_h$), which does not adjust for nonresponse, and Neyman allocation of expected respondents ($n_h^{\textrm{Ninv}}\propto N_h S_h / \bar{\phi}_h$), which inflates the sample size by the inverse of the anticipated response rates. In each case, we ignore the $S_h$ terms in our applications of these equations given our earlier assumption that $S_h = S$ for some $S$.

```{r}
nh_Ninv_propto <- pevs_adm_2016_rrs$Nhat_h
nh_Ninv <- nh_Ninv_propto / sum(nh_Ninv_propto) * 50000

nh_Nresp_propto <- pevs_adm_2016_rrs$Nhat_h / pevs_adm_2016_rrs$rr_h
nh_Nresp <- nh_Nresp_propto / sum(nh_Nresp_propto) * 50000

#Make sure allocations are valid
all(nh_Ninv >= 2 & nh_Ninv <= pevs_adm_2016_rrs$Nhat_h)
all(nh_Ninv >= 2 & nh_Nresp <= pevs_adm_2016_rrs$Nhat_h)
```

## View allocations

The results can be seen to match the tables shown in our paper's Appendix.

```{r}
(pevs_adm_allocs_50k <- 
  pevs_adm_alloc_50k_merged %>%
  mutate(n_h_optA = pevs_optA_alloc_50k,
         n_h_Ninv = nh_Ninv,
         n_h_Nresp = nh_Nresp) %>%
  select(c("h", "n_h_Ninv", "n_h_Nresp", "n_h_optA", "n_h_optE", "zeta_h_optE")))
```

## Evaluate performance

Further, we can roughly examine performance by computing the number of respondents, the design effect from weighting (Kish, 1992), and effective number of respondents under each allocation, replicating Table 2 of our paper. The results show that under the given assumptions, the proposed allocation has an effective sample size that is 25% higher than the Neyman-style allocations.

### Code

```{r}
#Calculate Kish's deff_w for poststratified estimator under nonresponse
calculate_deff_w <- function (N_h, r_h) {
  stopifnot(length(N_h) == length(r_h))
  stopifnot(all(r_h >= 1))
  stopifnot(all(r_h <= N_h))
  testthat::expect_equal(all(r_h <= N_h), TRUE)

  w_h <- N_h/r_h

  wbar <- sum(w_h * r_h)/sum(r_h)
  deff_w <- 1 + 1/sum(r_h) * sum((w_h - wbar)^2 * r_h)/wbar^2

  (deff_w)
}

#Summarize results from allocation under assumption that r_h = n_h * phibar_h
summarize_alloc <- function(N_h, phibar_h, n_h) {
  n_total <- sum(n_h)
  r_total <- sum(phibar_h * n_h)
  rr <- r_total / n_total
  deff_w <- calculate_deff_w(N_h = N_h, r_h = n_h * phibar_h)
  eff_resp <- r_total / deff_w
  c(n_total, r_total, rr, deff_w, eff_resp)
}

#Summarize results of each allocation
pevs_adm_alloc_summary_list <-
  lapply(2:5,
         function(col) summarize_alloc(N_h = pevs_adm_2016_rrs$Nhat_h,
                                       phibar_h = pevs_adm_2016_rrs$rr_h,
                                       n_h = unname(unlist(pevs_adm_allocs_50k[,col]))))

names(pevs_adm_alloc_summary_list) <- names(pevs_adm_allocs_50k[,2:5])

pevs_adm_alloc_summary_table <- pevs_adm_alloc_summary_list %>%
  bind_rows() %>%
  t()

colnames(pevs_adm_alloc_summary_table) <- c("n_total", "r_total", "rr", "deff_w", "eff_resp")
```

### Results

```{r}
tibble::as_tibble(pevs_adm_alloc_summary_table)
```

## References

Federal Voting Assistance Program. (2017a). 2015-2016 Active Duty Military dataset. Federal Voting Assistance Program, U.S. Department of Defense. <https://www.fvap.gov/uploads/FVAP/Surveys/PEV51601_Pub.dta> [Accessed 17 January 2023].

Federal Voting Assistance Program (2017b). Post-Election Voting Surveys: Active Duty Military: Technical report 2016. Technical report, Federal Voting Assistance Program, U.S. Department of Defense. <https://www.fvap.gov/uploads/FVAP/Reports/PEVS_ADM_TechReport_Final.pdf> [Accessed 17 January 2023].

Kish, L. (1992). Weighting for unequal Pi. *Journal of Official Statistics*, *8*(2), 183--200.
