---
title: "Sample allocation example: 2016 PEVS-ADM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sample allocation example: 2016 PEVS-ADM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Overview

The example below shows how to use the `samplingNR` package to allocate a sample of 50,000 invitees, using methods described in our paper (Mendelson & Elliott, in press) when assuming equivalent strata variances across strata (i.e., $S_h = S$ for some constant $S$). The example shows how to replicate our paper's Table 2, which compares the proposed allocation with common alternatives in an application to a survey of military personnel.

The stratification, population counts, and response rate information used in this example are included in our package as the data set `pevs_adm_2016_rrs`, structured as a [tibble](https://tibble.tidyverse.org/). These data are based on secondary analysis of public use data for the 2016 Post-Election Voting Survey of Active Duty Military (PEVS-ADM; Federal Assistance Voting Program, 2017a, 2017b), as described in our paper.

## Compute allocations

### Exact optimal allocation (iterative)

The function `opt_nh_nonresp()` calculates the exact (i.e., iteratively computed) version of the optimal allocation under nonresponse, as provided in our paper and summarized in the function documentation. We assume that stratification follows that of `pevs_adm_2016_rrs`, and specify the strata population sizes `N_h` and anticipated response propensities `phibar_h`. We have omitted the optional argument `S_h`, leading to use of the default assumption that $S_h$ is constant across strata. Note that by fixing the total number of invitees (`n_total`), we are implicitly assuming use of the *constant cost per invitee* scenario from our paper (i.e., per unit costs are assumed constant across strata and do not vary by response status).

```{r example}
library(samplingNR)
library(magrittr)
library(dplyr)

#Compute exact optimum allocation for PEVS-ADM 2016 data set for n = 50k
#Assumes tau = 1 by default
pevs_optE_alloc_50k <- opt_nh_nonresp(N_h = pevs_adm_2016_rrs$Nhat_h,
                                      phibar_h = pevs_adm_2016_rrs$rr_h,
                                      n_total = 50000)

#Merge results into original data frame
pevs_adm_alloc_50k_merged <- pevs_adm_2016_rrs %>%
  dplyr::mutate(n_h_optE = c(pevs_optE_alloc_50k),
                zeta_h_optE = attr(pevs_optE_alloc_50k,"zeta_h_prev"))

#View results
pevs_adm_alloc_50k_merged
```

### Approximately optimal allocation

The approximate version of this allocation, which assumes that $\zeta_h(n_h,\bar{\phi}_h) = 1$, can be computed using the function `opt_nh_nonresp_oneiter()`.

```{r}
pevs_optA_alloc_50k <- opt_nh_nonresp_oneiter(N_h = pevs_adm_2016_rrs$Nhat_h,
                                              phibar_h = pevs_adm_2016_rrs$rr_h,
                                              n_total = 50000)
```

### Neyman-style allocations for comparison

For comparison, we also compute Neyman allocation of invitees ($n_h^{\textrm{Ninv}}\propto N_h S_h$), which does not adjust for differential nonresponse, and Neyman allocation of expected respondents ($n_h^{\textrm{Nresp}}\propto N_h S_h / \bar{\phi}_h$), which adjusts by the inverse of the anticipated response rates. In each case, we ignore the $S_h$ terms in our applications of these equations, for purposes of this example, given our earlier assumption of equivalent strata variances.

```{r}
nh_Ninv_propto <- pevs_adm_2016_rrs$Nhat_h
nh_Ninv <- nh_Ninv_propto / sum(nh_Ninv_propto) * 50000

nh_Nresp_propto <- pevs_adm_2016_rrs$Nhat_h / pevs_adm_2016_rrs$rr_h
nh_Nresp <- nh_Nresp_propto / sum(nh_Nresp_propto) * 50000

#Make sure allocations are valid
all(nh_Ninv >= 2 & nh_Ninv <= pevs_adm_2016_rrs$Nhat_h)
all(nh_Nresp >= 2 & nh_Nresp <= pevs_adm_2016_rrs$Nhat_h)
```

## View allocations

The resulting allocations can be seen to match those shown in our paper's Appendix.

```{r}
(pevs_adm_allocs_50k <- 
  pevs_adm_alloc_50k_merged %>%
  mutate(n_h_optA = pevs_optA_alloc_50k,
         n_h_Ninv = nh_Ninv,
         n_h_Nresp = nh_Nresp) %>%
  select(c("h", "n_h_Ninv", "n_h_Nresp", "n_h_optA", "n_h_optE", "zeta_h_optE")))
```

## Evaluate performance

Further, we can roughly examine performance by computing the number of respondents, the design effect from weighting ($deff_w$; Kish, 1992), and the effective number of respondents under each allocation, replicating Table 2 of our paper. The results show that under the given assumptions, the proposed allocation has an effective sample size that is 25% higher than the Neyman-style allocations.

### Code

```{r}
#Calculate Kish's deff_w for poststratified estimator under nonresponse
calculate_deff_w <- function (N_h, r_h) {
  stopifnot(length(N_h) == length(r_h))
  stopifnot(all(r_h >= 1))
  stopifnot(all(r_h <= N_h))
  testthat::expect_equal(all(r_h <= N_h), TRUE)

  w_h <- N_h/r_h

  wbar <- sum(w_h * r_h)/sum(r_h)
  deff_w <- 1 + 1/sum(r_h) * sum((w_h - wbar)^2 * r_h)/wbar^2

  (deff_w)
}

#Summarize results from allocation under assumption that r_h = n_h * phibar_h
summarize_alloc <- function(N_h, phibar_h, n_h) {
  n_total <- sum(n_h)
  r_total <- sum(phibar_h * n_h)
  rr <- r_total / n_total
  deff_w <- calculate_deff_w(N_h = N_h, r_h = n_h * phibar_h)
  eff_resp <- r_total / deff_w
  c(n_total, r_total, rr, deff_w, eff_resp)
}

#Summarize results of each allocation
pevs_adm_alloc_summary_list <-
  lapply(2:5,
         function(col) summarize_alloc(N_h = pevs_adm_2016_rrs$Nhat_h,
                                       phibar_h = pevs_adm_2016_rrs$rr_h,
                                       n_h = unname(unlist(pevs_adm_allocs_50k[,col]))))

names(pevs_adm_alloc_summary_list) <- names(pevs_adm_allocs_50k[,2:5])

pevs_adm_alloc_summary_table <- pevs_adm_alloc_summary_list %>%
  bind_rows() %>%
  t()

colnames(pevs_adm_alloc_summary_table) <- c("n_total", "r_total", "rr", 
                                            "deff_w", "eff_resp")

```

### Results

```{r}
(pevs_adm_alloc_summary_tibble <- 
   tibble::as_tibble(pevs_adm_alloc_summary_table, rownames = "alloc"))

#View the percent increase in effective sample size from optE relative to Ninv 
# and Nresp.
eff_resp_vec <- pevs_adm_alloc_summary_tibble$eff_resp
eff_resp_vec[4]/eff_resp_vec[1:2] - 1
```

## References

Federal Voting Assistance Program. (2017a). 2015-2016 Active Duty Military dataset. Federal Voting Assistance Program, U.S. Department of Defense.

Federal Voting Assistance Program (2017b). Post-Election Voting Surveys: Active Duty Military: Technical report 2016. Technical report, Federal Voting Assistance Program, U.S. Department of Defense.

Kish, L. (1992). Weighting for unequal Pi. *Journal of Official Statistics*, *8*(2), 183--200.

Mendelson, J., & Elliott, M. R. (in press). Optimal allocation under anticipated nonresponse. *Journal of Survey Statistics and Methodology*.
